---
title: "Creating a kaggle-in-class challenge"
author: "Di Cook, Monash EBS"
output: 
    learnr::tutorial:
        progressive: true
        allow_skip: true
        css: "css/styles.css"
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  error = FALSE)
```

Image: [kowari](https://upload.wikimedia.org/wikipedia/commons/a/ac/Vakorejsek_ctyrprsty.jpg)

## Why?

Kaggle-in-class provides an objective way for students to assess their learning of predictive modeling. 

Examples from recent years:

- 2020: [Pictionary](https://www.kaggle.com/c/pictionary/leaderboard)
- 2019: [How Points End in Tennis](https://www.kaggle.com/c/monba2019)
- 2017: [Melbourne auction prices](https://www.kaggle.com/c/vitticeps/leaderboard)
- 2017: [Spam or Ham](https://www.kaggle.com/c/spam-or-not-spam/leaderboard)
- 2016: [Ames housing prices](https://www.kaggle.com/c/galah/leaderboard)
- 2016: [Who's speaking](https://www.kaggle.com/c/cockatoo/leaderboard)
- 2015: [Happy paintings](https://www.kaggle.com/c/monba/leaderboard)

## Getting data ready

### Choosing data

- the full set is not available to the students, to avoid plagiarism and use of unauthorized assistance.
- the data is not too easy, or too hard, to model so that there is some discriminatory power in the results.
- data should be relatively clean, to the point where the instructor has tested that a model can be fitted.
- contains some challenges, that make standard off-the-shelf modeling less successful, like different variable types that need processing or transforming, some outliers, large number of variables.
- if it is a classification challenge, it will work better with relatively balanced classes, because the overall accuracy is the easiest metric to use.

### Let's get wildlife sightings around Monash

I've pulled the occurrence records from https://www.ala.org.au. 

```{r  eval=FALSE}
monash <- read_csv("monash/monash.csv")
counts <- monash %>% count(`Vernacular name`, sort=TRUE)
keep <- c("Rainbow Lorikeet", "Galah", "Dusky Moorhen", "Umbrella tree", "Spotted Pardalote")
monash_ala <- monash %>% filter(`Vernacular name` %in% keep) %>%
  select(`Vernacular name`, Longitude, Latitude) %>%
  rename(name=`Vernacular name`)
save(monash_ala, file="monash/monash_ala.rda")
```

```{r}
library(tidyverse)
library(ggmap)
load("monash/monash_map.rda")
load("monash/monash_ala.rda")
ggmap(map) + 
  geom_point(data=monash_ala, aes(x=Longitude, y=Latitude, colour=name, shape=name), size=2, alpha=0.7) +
  scale_colour_brewer(palette="Dark2")
```

### Generate training and test sets

Using the `tidymodels` approach, break the data into a training and testing set. Use 50% because kaggle will further break the data.

```{r echo=TRUE}
library(tidymodels)
set.seed(444)
data_split <- initial_split(monash_ala, prop = 1/2)
monash_ala_tr <- training(data_split)
monash_ala_ts <- testing(data_split)
```

### Checks

Check that it properly stratified the classes.

```{r echo=TRUE, eval=FALSE}
monash_ala %>% count(name)
monash_ala_tr %>% count(name)
monash_ala_ts %>% count(name)
```

Check that the order of classes is a bit mixed

```{r echo=TRUE, eval=FALSE}
monash_ala_tr$name
monash_ala_tr$name
```

### Create unlabelled test set, solution and example solution data sets

Add an id column, so that when labels are removed, observations can be matched back to original set, or matched against solution set. Also the latest version of kaggle-in-class requires the important columns be named `Id` and `Category`. 

```{r echo=TRUE}
monash_ala_ts <- monash_ala_ts %>% 
  mutate(Id = 1:n()) %>%
  rename(Category = name) %>%
  select(Id, Category, Longitude, Latitude)
monash_ala_tr <- monash_ala_tr %>%
  rename(Category = name)
```


Need to provide an unlabelled test set for competitors. Save the full test set into a new file, clearly labelled and remove the labels in the test object. 

```{r echo=TRUE}
monash_ala_ts_labelled <- monash_ala_ts
monash_ala_ts <- monash_ala_ts %>% 
  mutate(Category = NA)
monash_ala_ts %>% print(n=5)
```

Create a solution set, and a sandbox (sample solution) set. 

```{r echo=TRUE}
monash_ala_solution <- monash_ala_ts_labelled %>%
  select(Id, Category)
monash_ala_example_solution <- monash_ala_solution %>%
  mutate(Category = sample(Category))
```

Write to files for kaggle upload.

```{r echo=TRUE}
write_csv(monash_ala_tr, path="kaggle_files/monash_ala_tr.csv")
write_csv(monash_ala_ts, path="kaggle_files/monash_ala_ts.csv")
write_csv(monash_ala_solution, path="kaggle_files/monash_ala_solution.csv")
write_csv(monash_ala_example_solution, path="kaggle_files/monash_ala_example_solution.csv")
```

## Check modeling

Its a good idea to provide a basic model as a helper for students to get started.

```{r eval=FALSE}
ct_spec <- decision_tree() %>% 
  set_engine("rpart") %>% 
  set_mode("classification")
monash_ala_rec <- recipe(Category~., data=monash_ala_tr)
monash_ala_fit <- 
  workflow() %>% 
  add_model(ct_spec) %>% 
  add_recipe(monash_ala_rec) %>%
  fit(data=monash_ala_tr)
```

```{r echo=TRUE}
library(rpart)
monash_ala_rp <- rpart(Category~., data=monash_ala_tr,
              control = rpart.control(cp = 0.0005, minsplit = 5)) 
# monash_ala_rp

monash_ala_ts_pred <- predict(monash_ala_rp, monash_ala_ts, type="class")
table( monash_ala_ts_labelled$Category, monash_ala_ts_pred)
```

### Provide guidelines to students

For the pictionary challenge, [these](https://iml.numbat.space/project/project.html) were the instructions.

## Uploading to kaggle

### Finding the kaggle-in-class 

- Go to the main kaggle host a competition site: https://www.kaggle.com/c/about/host/
- Look for link `Interested in hosting a classroom competition? Visit Kaggle InClass »`
- This takes you to page on help and to create a competition: https://www.kaggle.com/c/about/inclass

### Create a competition

1. Click `Create competition` link
2. Give a short name and description
3. Fill in overview

### Competition details

4. Choose evaluation metric
5. Provide information on submission file format
6. Upload data
7. Add descriptions of the data
8. Set the rules
9. Set deadlines - careful with dates and times
10. Teams or no teams, number of submissions per day allowed

### Styling

11. Add images

### Access

12. Limit to monash email address? Or to anyone with link?

### Add solution set

13. Upload solution file
14. Set public/private split
15. Upload sample submission file

### Try it out

15. Sandbox solutions

### Go live

16. Launch checklist
17. Launch competition now

### Test it out!

This is your turn!

- Download the training data and test data. 
- Make a prediction of the test set.
- Upload your submission. 

## Resources

- Kaggle site instructions: https://www.kaggle.com/about/inclass/overview
- Julia Polak's video instructions: 
